Analyzing Information Transfer from Different Data Sources via Precise Bias-variance Asymptotics

### 2021-07-05 revision_journal/intro.tex
Updates from Overleaf

In the single task case, if the data is Gaussian distributed, then precise variance can be derived immediately from properties of Wishart distribution. In particular, in the high dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}. It well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. 
In the single-task case, \cite{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\cite{Anisotropic} later extended this result to sample covariance matrices with arbitrary covariance.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, a recent development in the random matrix theory literature.
These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for two tasks (with arbitrary covariate shifts).
%\HZ{can you please add a discussion of the related work from RMT?}

>>>>

These previous works mainly consider cases where there is only one high-dimensional sample covariance matrix involved. ......... In the single task case, if the data is Gaussian distributed, then precise variance can be derived immediately from properties of Wishart distribution. In particular, in the high dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}. It well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. 
In the single-task case, \cite{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\cite{Anisotropic} later extended this result to sample covariance matrices with arbitrary covariance.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, a recent development in the random matrix theory literature. 
%These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
%We also extend the techniques in ?, which is one contribution of this work.

There exists another method to derive the asymptotic variance limit using free addition in free probability theory \cite{nica2006lectures}. We do not take this procedure because this method is not completely justified in the general non-Gaussian setting (in the sense that it is hard to establish rigorously the free independence between two sample covariance matrices with non-diagonal covariate shift). Furthermore, our techniques provide almost sharp convergence rates to the asymptotic limit, which cannot be given by free probability theory. To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for two tasks (with arbitrary covariate shifts).

On the other hand, the bias limit is much harder to derive. However, when the data is Gaussian distributed and there is no covariante shift, we can use recent results in ... to derive the bias limit.

%\HZ{can you please add a discussion of the related work from RMT?}

### 2021-07-05 revision_journal/results.tex
Updates from Overleaf

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/covariate_shift.eps}
		\caption{Example \ref{ex_covshift}}
		\label{fig_covariate}
	\end{subfigure}
\end{figure}

\begin{example}[Whether or not covariate shift helps depends on sample sizes]
    %However, we now use an example to show that the above observation fails when $n_1$ is comparable to $n_2$.
    Our first example illustrates that the effect of covariate shift depends on the sample sizes of each dataset in an intricate manner.
    Let $\lambda > 0$ be a variable that captures the degree of covariate shift: higher values imply worse covariate shift.
    Suppose that half of the singular values of $M$ are equal to $\lambda$ while the other half eigenvalues are $\lambda_{p/2}^{-1}\ge \cdots \ge \lambda_2^{-1} \ge \lambda_1^{-1}$. When $\lambda_1/\lambda_{p/2} = 1$, we have $M=\id_{p\times p}$ and there is no covariate shift.
%while as As $\lambda$ increases, the severity of covariate shift increases.
We claim the following dichotomy.
\begin{enumerate}
	\item If $n_1 \ge n_2$, then the variance limit is smallest when there is no covariate shift.
	\item If $n_1 < n_2$, then the variance limit is largest when there is no covariate shift.
\end{enumerate}
We explain why the above dichotomy happens. We can write the variance limit as
$$\ell_{\var}(M)=\frac{\sigma^2}{ n_1+n_2 }\sum_{i=1}^{p/2}\left( \frac{1}{\lambda_i^{2} a_1 + a_2} + \frac1{\lambda_i^{-2} a_1 + a_2} \right) .$$
When $M=\id_{p\times p}$, by the first equation of \eqref{eq_a12extra}, we have
$$\ell_{\var}(\id_{p\times p})=\frac{\sigma^2}{ n_1+n_2 }\sum_{i=1}^{p/2} \frac{2}{1-\gamma} , $$
where we abbreviate $\gamma:=p/(n_1+n_2)$. Then using $a_1+a_2=1-\gamma$, through a direct calculation we find that %we can calculate that
\begin{align*}
\ell_{\var}(M) - \ell_{\var}(\id_{p\times p})%&= \frac{\sigma^2 \gamma}{2(1-\gamma)} (\lambda^2-1)a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{\sigma^2 }{ n_1+n_2-p} \sum_{i=1}^{p/2} \frac{(\lambda_i^2-1)^2 a_1(M) \left[ a_1(M) - a_2(M)\right] }{[a_1(M)+\lambda_i^2 a_2(M)][\lambda_i^2 a_1(M)+a_2(M)]} .
\end{align*}
%in this example is equal to
%$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
%\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
%Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
%\begin{align*}
%	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
%	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
%\begin{align*}
%	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
%where $g(\lambda, a_1) \ge 0$.
%and can be derived from algebraic calculations (details omitted).
We claim that $a_1(M) > a_2 (M)$ if and only if $n_1 > n_2$, which then explains the dichonomy. In fact, if $a_1>a_2$, then the first equation in \eqref{eq_a12extra} gives that $a_1> (1-\gamma)/2$, and the second equation in  \eqref{eq_a12extra} gives that
\begin{align*}
 \frac{n_1}{n_1 + n_2} &> a_1 + \frac{1}{n_1+n_2} \sum_{i=1}^{p/2}\left(\frac{\lambda_i^2}{\lambda_i^2+1}+\frac{\lambda_i^{-2}}{\lambda_i^{-2}+1}\right) = \frac{1-\gamma}{2}+\frac{\gamma}{2}=\frac{1}{2}.
\end{align*}
%\begin{align*}
% \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
%	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
%\end{align*}
This implies $n_1>n_2$. The other direction follows from a similar argument. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$. Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.
%\end{example}

\todo{draw a connection to \citet{david2010impossibility}}
\end{example}

%\paragraph{Illustrative examples.}
%Hence we take $a=1$ in \eqref{lem_cov_shift_eq}, and study the asymptotic limit
%$$\ell_{\var}(M):= \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1(M) \cdot M^\top M + a_2(M)  }  }, $$
%where $ M=(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and $a_1(M)$ and $a_2(M)$ are defined through \eqref{eq_a12extra}.
\begin{example}[Covariate shift hurts more under imbalanced sample sizes]
We compare different choices of $M$ that are scaled to have the same determinant. More precisely, for a fixed $\mu>0$ we define
		\cS_{\mu}\define\bigset{M \left| \prod_{i=1}^p \lambda_i = \mu^p, \tau \le \lambda_p \le \lambda_1 \le \tau^{-1}\right.}.
%belong to the following bounded set.
%Let $\lambda_i$ be the $i$-th singular value of $M$.
%let $\mu_{\min} < \mu < \mu_{\max}$ be fixed values that do not grow with $p$.
%\vspace{-0.025in}
%{\small}
%	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
%\begin{proposition}[Covariate shift]\label{prop_covariate}
%	Assume that $\Psi(\beta_1, \beta_2) = 0$ and $\rho_1, \rho_2>1$.
%	Let $g(M)$ denote the prediction loss of $\hat{\beta}_t^{\MTL}$ when $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} \in\cS_{\mu}$.
%	We have that
%	{\small\[ g(\mu\id) \le \bigbrace{1+ \bigo{{\rho_2}/{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]}
%\end{proposition}
%This proposition shows that when source/target sample ratio is large, then having no covariate shift is optimal.
%The proof of Proposition \ref{prop_covariate} is left to Appendix \ref{app_proof_33}.
%We now prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix where there is enough source data.
Our first observation is that if $n_1 \gg n_2$, i.e. there is enough source data compared to the target data, then $\ell_{\var}(M)$ is minimized approximately when $M$ is a scalar matrix.
%\begin{proof}[Proof of Proposition \ref{prop_covariate}]
%Let
%$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$
%We now calculate $g(M_0)$. With the same arguments as in Lemma \ref{lem_hat_v}, we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2$ as in \eqref{choiceofpara}, we can simplify
%\be \nonumber
%\begin{split}
%g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  ,
%\end{split}
%\ee
%with high probability for some constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $ \mu_{\min}\le \lambda_p(M)\le \lambda_1(M)\le \mu_{\max}$. We omit the details here.
%
%Now using Lemma \ref{lem_cov_shift}, we obtain that with high probability,
%\begin{align}\label{gvar_extra}
%g(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1(M_0)\cdot M_0^\top M_0 + a_2(M_0)}\right) \cdot \left(1 +\OO(p^{-\e})\right).
%\end{align}
In fact, from equation \eqref{eq_a12extra}, we obtain the following estimates on $ a_1(M)$ and $a_2(M)$ for any $M\in \cal S_\mu$:
\be\label{est_a12extra}
 a_1(M) = \frac{n_1+\OO(p)}{n_1+n_2},\quad a_2(M) =\frac{n_2+\OO(p)}{n_1+n_2}.
Inserting \eqref{est_a12extra} into $\ell_{\var}(M)$, we obtain that
%and using $ M_0^\top M_0\succeq \mu_{\min}^2$, we obtain that with high probability,
%\begin{align}\label{approximateteM}
%\left(1+\frac{\rho_2}{(\rho_1-1)\mu_{\min}^2}\right)^{-1}h(M_0) \cdot \left(1 - \OO(p^{-\e})\right) \le g(M_0) \le h(M_0) \cdot \left(1 +\OO(p^{-\e})\right),
%\end{align}
%where
$$\ell_{\var}(M)= \left[ 1+\OO\left(\frac{n_2}{n_1}\right)\right] \cdot \frac{\sigma^2}{n_1} \tr\left( \frac{1}{M^\top M}\right) .$$
%With these two bounds, we can easily conclude \eqref{approxteM}.
%
%We have that the test error satisfies
%\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
%where $\lambda_p$ is the smallest singular value of $p$ and
%$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
%Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.
By AM-GM inequality, we observe that
$$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i^2}$$
is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i =\mu^p$. Hence in this case, $M=\mu \id_{p\times p}$ is approximately the optimal choice.
\end{example}
%Hence we get that
%\be\label{AMGM} h(M_0) \le \frac{\sigma^2}{\mu^2 (\rho_1+\rho_2)a_1(M_0)}.\ee
%
%On the other hand, when $M=\mu \id$, applying Lemma \ref{lem_cov_shift} we obtain that with high probability,
%\begin{align}\label{gvar_extra2}
%\begin{split}
%g(\mu \id)&= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{\mu^2 a_1 (\mu\id) + a_2(\mu\id)}\right) \cdot \left(1 +\OO(p^{-\e})\right)\\
%&\le \frac{\sigma^2}{\mu^2(\rho_1+\rho_2)a_1 (\mu\id)}.
%\end{split}
%\end{align}
%Combining \eqref{est_a12extra}, \eqref{approximateteM}, \eqref{AMGM} and \eqref{gvar_extra2}, we conclude the proof.
%, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
%\end{proof}

\iffalse
For large enough $p$,
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
\begin{align*}
	\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} &\rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}
	= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}.
\end{align*}
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit depends on the spectrum of $M$. % To be clear, for this example we assume that the bias is $0$.
\fi

%Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}.
%Our first observation is that:

\iffalse
%FY: Below was the word explanation. I still think the above simple proof is clearer.
In fact, due to the fact that $M=M^{-1}$, $a_1$ and $a_2$ play symmetric roles in equations \eqref{eq_a12extra000} and \eqref{eq_a12extra}.
Hence, when $n_1 \ge n_2$, we have that $a_1 \ge a_2$, hence $a_1 \ge \frac{1}{2}(1 - \frac{p}{n_1 + n_2 - p}) = \frac{n_1 + n_2}{2 (n_1 + n_2 - p)}$.
The other case when $n_1 < n_2$ is similar.
A formal proof follows easily from the self-consistent equations \eqref{lem_cov_shift_eq} and we omit the details.
Thus, we conclude that if $n_1 \ge n_2$, then $f(\lambda) > f(1)$.
If $n_1< n_2$, then $f(\lambda)< f(1)$.
\fi

\paragraph{Comparing HPS to other estimators.}

>>>>

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/covariate_shift.eps}
		\caption{HPS estimator under covariate shift}
		\label{fig_sec3_covariate}
	\end{subfigure}
	\caption{The dimension $p = 200$, $n_2 = 800$, and the noise variance $\sigma^2 = 1/4$.}
\end{figure}

\subsubsection{Illustrative examples}

@@

\begin{example}[Does covariate shift help or hurt information transfer?]
    %However, we now use an example to show that the above observation fails when $n_1$ is comparable to $n_2$.
    Our first example illustrates that the effect of covariate shift depends on the sample sizes of each dataset in an intricate manner.
    While the folklore belief is that transferring from a covariate-shifted distribution performs worse than an identical distribution, we show that, surprisingly, the former can sometimes outperform the latter.

    We first describe a setting for modeling covariate shift.
    Let $\cS$ be the set of positive definite matrices $M^{\top} M$ such that for $i = 1,2,\dots,\floor{\frac p 2}$, $\lambda_{p - i}  = \frac 1 {\lambda_i}$, where $\lambda_1 \ge \lambda_2 \ge \cdots \ge \lambda_p$ are the eigenvalues of $M$.
    Under this setting, we ask: suppose that the sample sizes $n_1$ and $n_2$ are both fixed, which $M$ provides the best ``transfer'' in the sense of HPS estimator's excess risk?
    We observe a dichotomy that depends on whether or not $n_1$ is greater than $n_2$.
    \begin{claim}\label{claim_dichotomy}
        Let $g(M) = \frac{\sigma^2}{n_1 + n_2} \bigtr{\frac 1 {\alpha_1 M^{\top} M + \alpha_2}}$.
        Suppose that $n_1$ and $n_2$ are both fixed.
        Within the set of all possible covariate shift matrix $M$ described above, the following dichotomy holds:
        \begin{enumerate}
	        \item If $n_1 \ge n_2$, then $g(M)$ is minimized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
	        \item If $n_1 < n_2$, then $g(M)$ is maximized in $\cS$ when $\lambda_i = 1$, for any $i = 1,\dots,p$.
        \end{enumerate}
    \end{claim}

\begin{proof}
    We explain why the above dichotomy happens. We can write $g(M)$ for any $M \in \cS$ as
    $$g(M)=\frac{\sigma^2}{ n_1+n_2 }\sum_{i=1}^{p/2}\left( \frac{1}{\lambda_i^{2} a_1 + a_2} + \frac1{\lambda_i^{-2} a_1 + a_2} \right).$$
    When $M=\id_{p\times p}$, by the first equation of \eqref{eq_a12extra}, we have
    $$g(\id_{p\times p})=\frac{\sigma^2}{ n_1+n_2 }\sum_{i=1}^{p/2} \frac{2}{1-\gamma},$$
    where we abbreviate $\gamma:=p/(n_1+n_2)$. Then using $\alpha_1 + \alpha_2 = 1-\gamma$, through a direct calculation we find that %we can calculate that
    \begin{align*}
        g(M) - g(\id_{p\times p})%&= \frac{\sigma^2 \gamma}{2(1-\gamma)} (\lambda^2-1)a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
        &= \frac{\sigma^2 }{ n_1+n_2-p} \sum_{i=1}^{p/2} \frac{(\lambda_i^2-1)^2 \alpha_1 \left[ \alpha_1 - \alpha_2\right] }{[\alpha_1 + \lambda_i^2 \alpha_2][\lambda_i^2 \alpha_1 + \alpha_2]} .
    \end{align*}
    %in this example is equal to
    %$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
    %\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
    %Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
    %\begin{align*}
    %	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
    %	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
    %	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
    %	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
    %\end{align*}
    %\begin{align*}
    %	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
    %\end{align*}
    %where $g(\lambda, a_1) \ge 0$.
    %and can be derived from algebraic calculations (details omitted).
    We claim that $\alpha_1 > \alpha_2$ if and only if $n_1 > n_2$, which then explains the dichotomy.
    In fact, if $\alpha_1 > \alpha_2$, then the first part of equation \eqref{eq_a12extra} gives that $\alpha_1 > (1-\gamma)/2$.
    The second part of equation \eqref{eq_a12extra} gives that
    \begin{align*}
     \frac{n_1}{n_1 + n_2} &> \alpha_1 + \frac{1}{n_1+n_2} \sum_{i=1}^{p/2}\left(\frac{\lambda_i^2}{\lambda_i^2+1}+\frac{\lambda_i^{-2}}{\lambda_i^{-2}+1}\right) = \frac{1-\gamma}{2}+\frac{\gamma}{2}=\frac{1}{2}.
    \end{align*}
    %\begin{align*}
    % \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
    %	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
    %\end{align*}
    This is equivalent to $n_1>n_2$. The other direction follows from a similar argument. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$. Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.
    %\end{example}
\end{proof}

Figure \ref{fig_sec3_covariate} illustrates a special case where $\lambda_1 = \cdots = \lambda_{\floor{p /2}} = \lambda > 0$ for $M^{\top}M$ and the rest of the eigenvalues are all equal to $1 / \lambda$.
Thus, $\lambda > 0$ captures the degree of covariate shift: higher $\lambda$ implies worse covariate shift.
We observe that our theoretical estimates using Corollary \ref{cor_hps_cov} matches the empirical risks incredibly accurately.
As a result, we indeed observe the dichotomy shown in Claim \ref{claim_dichotomy}.
\end{example}

%\paragraph{Illustrative examples.}
%Hence we take $a=1$ in \eqref{lem_cov_shift_eq}, and study the asymptotic limit
%$$\ell_{\var}(M):= \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1(M) \cdot M^\top M + a_2(M)  }  }, $$
%where $ M=(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and $a_1(M)$ and $a_2(M)$ are defined through \eqref{eq_a12extra}.
\begin{example}[Covariate shift hurts more under imbalanced sample sizes]
Our second example illustrates that the effect of covariate shift worsens as the sample size of task one increases.
We consider a set of covariate shift matrices $M^{\top} M$ whose determinant are all equal to one, and the eigenvalues are bounded between $\tau^2$ and $1 / \tau^2$.
Let $\cS$ denote the set of $M\in\real^{p\times p}$ such that $M^{\top} M$ satisfies both conditions.
%\begin{align*}
%		\cS_{\mu}\define\bigset{M \left| \prod_{i=1}^p \lambda_i = \mu^p, \tau \le \lambda_p \le \lambda_1 \le \tau^{-1}\right.}.
%\end{align*}
\begin{claim}\label{prop_covariate}
%	Assume that $\Psi(\beta_1, \beta_2) = 0$ and $\rho_1, \rho_2>1$.
%	Let $g(M)$ denote the prediction loss of $\hat{\beta}_t^{\MTL}$ when $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} \in\cS_{\mu}$.
    Recall that $g(M) = \frac{\sigma^2}{n_1 + n_2} \bigtr{\frac 1 {\alpha_1 M^{\top} M + \alpha_2}}$.
	We have the following relation between $g(\id_{p \times p})$ and the rest of $g(M)$ for any $M \in\cS$:
	\[ g(\id_{p\times p}) \le \bigbrace{1+ \bigo{\frac{n_2}{n_1}  }} g(M), \text{ for any } M \in \cS. \]
\end{claim}
\begin{proof}
    From equation \eqref{eq_a12extra}, it is easy to obtain the following estimates on $\alpha_1$ and $\alpha_2$ for any $M\in \cS$:
    \begin{align*}%\label{est_a12extra}
        \frac{n_1 - p}{n_1 + n_2} < \alpha_1 <  \frac{n_1 + n_2 - p}{n_1 + n_2},\quad \alpha_2 < \frac{n_2}{n_1 + n_2}.
    \end{align*}
    Applying the above into $g(M)$, and using the fact that the minimum eigenvalue of $ M^\top M$ is at least $\tau^2$, we obtain
    \begin{align}\label{approximateteM}
    \left(1+\frac{\rho_2}{(\rho_1-1)\mu_{\min}^2}\right)^{-1}h(M_0) \cdot \left(1 - \OO(p^{-\e})\right) \le g(M_0) \le h(M_0) \cdot \left(1 +\OO(p^{-\e})\right),
    \end{align}
    where
    $$h(M_0):=\frac{\sigma^2}{(\rho_1+\rho_2)a_1(M_0)}\cdot \frac1p\tr\left( \frac{1}{M_0^\top M_0}\right) .$$
    %With these two bounds, we can easily conclude \eqref{approxteM}. 
    %
    %We have that the test error satisfies
    %\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
    %where $\lambda_p$ is the smallest singular value of $p$ and
    %$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
    %Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.
    By AM-GM inequality, we observe that 
    $$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i^2}$$
    is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i\le \mu^p$. Hence we get that 
    \be\label{AMGM} h(M_0) \le \frac{\sigma^2}{\mu^2 (\rho_1+\rho_2)a_1(M_0)}.\ee

    On the other hand, when $M=\mu \id$, applying Lemma \ref{lem_cov_shift} we obtain that with high probability,
    \begin{align}\label{gvar_extra2}
    \begin{split}
    g(\mu \id)&= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{\mu^2 a_1 (\mu\id) + a_2(\mu\id)}\right) \cdot \left(1 +\OO(p^{-\e})\right)\\
    &\le \frac{\sigma^2}{\mu^2(\rho_1+\rho_2)a_1 (\mu\id)}.
    \end{split}
    \end{align}
    Combining \eqref{est_a12extra}, \eqref{approximateteM}, \eqref{AMGM} and \eqref{gvar_extra2}, we conclude the proof.
    %, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
\end{proof}
\todo{draw a connection to \citet{david2010impossibility}}
\end{example}

\subsubsection{Comparing HPS to other estimators}
